\documentclass[oneside]{book}
\usepackage{manmac}
%\usepackage[toc,page]{appendix}
%\pagenumbering{arabic}

\title{The Brainix Manual}

\author{The Brainix Team  \\
{\small\em Copyleft \copyright \ Draft date \today }}

\date{ }
\begin{document}
\frontmatter


\maketitle
%\addcontentsline{toc}{chapter}{Contents}
%\pagenumbering{roman}
\tableofcontents

%\mainmatter
\part{Introductory matters}
\include{tex/todo}
\mainmatter
\part{Understanding the Brainix Operating System}
\include{tex/fs}
%\include{tex/vfs}
\part{Tutorials on Using the Brainix Operating System}
\include{tex/run}
\include{tex/hack}
\include{tex/debug}


\part{Appendices}
\appendix
\include{tex/fdl}
\backmatter 
%\renewcommand{\refname}{\chapter{Bibliography}}
\addcontentsline{toc}{chapter}{Bibliography}
\begin{thebibliography}{10}
\bibitem{skelix} The Skelix Operating System Introduction to writing (an operating system including) a file system from scratch \url{http://en.skelix.org/skelixos/tutorial07.php}
\bibitem{minixbook} Andrew Tanenbaum and Albert Woodhull, \textit{Operating Systems: Design and Implementation} Third Edition, Upper Saddle River, New Jersey: Pearson Prentice Hall (2006).
\bibitem{linuxbook} Daniel P. Bovet and Marco Cesati, \textit{Understanding the Linux Kernel} Sebastopol: O'Reilly Media Inc. (2006).
\bibitem{bsdbook} Marshall Kirk McKusick and George V. Neville-Neil \textit{The Design and Implementation of the FreeBSD Operating System: FreeBSD Version 5.2} Upper Saddle River: Pearson Education, Inc. (2005).
\bibitem{1} Dr. Matloff's ``File Systems in Unix'' \url{http://heather.cs.ucdavis.edu/~matloff/UnixAndC/Unix/FileSyst.html} 
\bibitem{2} Unix File System \url{http://unixhelp.ed.ac.uk/concepts/fsystem.html}
\bibitem{3} The Linux File System Explained, by Mayank Sarup \url{http://www.freeos.com/articles/3102/}
\bibitem{4} Chapter 9 of \textit{The Linux Kernel} Available \url{http://www.tldp.org/LDP/tlk/fs/filesystem.html}
\bibitem{5} Inode Definition by the Linux Information Project \url{http://www.bellevuelinux.org/inode.html}
\bibitem{6} Understanding the Filesystem Inodes \url{http://www.onlamp.com/pub/a/bsd/2001/03/07/FreeBSD_Basics.html}
\bibitem{7} David A Rusling, The Second Extended File system (EXT2) \url{http://www.science.unitn.it/~fiorella/guidelinux/tlk/node97.html}
\bibitem{8} Understanding the Linux Kernel Second Edition Excerpt \url{www.oreilly.com/catalog/linuxkernel2/chapter/ch17.pdf}
\bibitem{9} ``Setting your file creation mask'' \url{http://osr507doc.sco.com/en/OSTut/Setting_your_file_creation_mask.html}
\bibitem{10} Kevin T. Van Maren \textit{The Fluke Device Driver Framework} (1999). \url{http://www.cs.utah.edu/flux/papers/vanmaren-thesis.pdf}
\bibitem{11} David A Rusling \textit{The Linux Kernel} Chapter 9 \url{http://www.tldp.org/LDP/tlk/fs/filesystem.html}
\bibitem{12} Balazs Gerofi \textit{MINIX VFS: Design and implementation of the MINIX Virtual File system} \url{http://www.minix3.org/doc/gerofi_thesis.pdf}
\bibitem{13} S.R. Kleiman \textit{Vnodes: An Architecture for Multiple File System Types in Sun UNIX} In Proc. of the Summer 1986 USENIX Technical Conference, June 1991, pp 238-247 \url{http://www.arl.wustl.edu/~fredk/Courses/cs523/fall01/Papers/kleiman86vnodes.pdf}
\bibitem{14} Neil Brown et al. \textit{The Linux Virtual File-system Layer} (1999) \url{http://www.cse.unsw.edu.au/~neilb/oss/linux-commentary/vfs.html}
\bibitem{15} Richard Gooch et al. \textit{Overview of the Linux Virtual File System} (2005) \url{http://www.gelato.unsw.edu.au/lxr/source/Documentation/kernel-docs.txt} 
\bibitem{16} Max Bruning \textit{A Comparison of Solaris, Linux, and FreeBSD Kernels} (2005) \url{http://www.opensolaris.org/os/article/2005-10-14_a_comparison_of_solaris__linux__and_freebsd_kernels/}
\bibitem{17} Vans Information \textit{Virtual file system Part 1} (2001) \url{http://freeos.com/articles/3838/}
\bibitem{18} Vans Information \textit{Virtual file system Part 2} (2001) \url{http://freeos.com/articles/3851/}
\bibitem{19} \url{http://fxr.watson.org/fxr/source/?v=RELENG62}
\bibitem{20} \url{http://fxr.watson.org/fxr/source/sys/?v=RELENG62}
\bibitem{21} Remy Card, Theodore Ts'o, and Stephen Tweedie, \textit{The Design and Implementation of the Ext2 File System} (1994) \url{http://web.mit.edu/tytso/www/linux/ext2intro.html}
\bibitem{22} Anon. \textit{The Ext2 Linux Documentation} (for kernel 2.6.20.1)  \url{http://lxr.linux.no/source/Documentation/filesystems/ext2.txt}
\end{thebibliography}
\end{document}
